---
title: "Study 3 Analysis"
author: "Liam A. Doyle"
date: "2024-05-03"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
library(tidyverse)
library(tidyr)
library(readxl)
library(psych)
library(gt)
library(corrplot)
library(ggcorrplot)
library(WRS2)
library(interactions)
library(jtools)
library(MASS)
library(lme4)
library(lmerTest)
library(sandwich)
library(clubSandwich)
library(sjPlot)
library(ggeffects)
library(marginaleffects)
```

## Loading Data

```{r load, include=FALSE}
data <- read_csv("analysis_wide.csv")
```

## Demographics of Sample

To get a basic picture of the sample being used for analysis, descriptives were run for common demographic variables (e.g., age, gender).

```{r demographic, echo=FALSE}

# Examining basic descriptives for age

describe(data$Age)

# Create histogram of age
ggplot(data, aes(x = Age)) +
  geom_histogram(breaks = seq(0.5, 42.5, by = 1),  # Set breaks at every year from 1 to 42
                 color = "black",                  # Border color of the bins
                 fill = "white") +                 # Fill color of the bins
  labs(title = "Distribution of Age",              # Main title
       x = "Age (years)",                          # X-axis label
       y = "Frequency") +                          # Y-axis label
  theme_minimal() +                                # Minimalist theme for a professional look
  theme(text = element_text(size = 12),            # Set text size
        plot.title = element_text(hjust = 0.5),    # Center the plot title
        axis.title = element_text(size = 14),      # Set axis titles size
        axis.text = element_text(size = 12)) 

# Examining basic descriptives for race

summary_df <- data %>%
  group_by(Race) %>%
  summarise(
    Frequency = n(),
    Percentage = n() / nrow(data) * 100
  ) %>%
  arrange(desc(Percentage))

print(summary_df)

# Examining basic descriptives for gender

summary_df <- data %>%
  group_by(Gender) %>%
  summarise(
    Frequency = n(),
    Percentage = n() / nrow(data) * 100
  )

print(summary_df)

# Examining basic descriptives for sex

summary_df <- data %>%
  group_by(Sex) %>%
  summarise(
    Frequency = n(),
    Percentage = n() / nrow(data) * 100
  )

print(summary_df)

```

## Creating a Table with Descriptive Statistics for Important Variables in the Study

An APA 7 style table was created with the descriptive statistics for important variables in the study (e.g., OS_Avg).

```{r descriptive key variables, echo=FALSE}

descriptive_stats <- describe(data[, c("OS_Avg", "OSF_Avg", "Diff_OS_Avg", 
                                       "IT_Avg", "ITF_Avg", "Diff_IT_Avg",
                                       "SRP_Total", "IES_Avg", "BZSG_Avg")])

print(descriptive_stats)
```

## Correlation Matrix Between Major Variables

Self-explanatory. Correlations between important variables in the study (i.e., variables relevant to hypothesis testing).

```{r correlation matrix, echo=FALSE}

# Selecting relevant variables
selected_data <- data[, c("OS_Avg", "OSF_Avg", "Diff_OS_Avg", 
                          "IT_Avg", "ITF_Avg", "Diff_IT_Avg",
                          "SRP_Total", "IES_Avg")]

# Compute the correlation matrix
cor_matrix <- cor(selected_data, use = "pairwise.complete.obs")

# Print the correlation matrix
print(cor_matrix)

# Plot the correlation matrix using corrplot
corrplot(cor_matrix, method = "color", type = "upper", order = "hclust",
         tl_col = "black", tl_srt = 45)  # Adjust text label color and rotation for better readability

# Plot the correlation matrix using ggcorrplot

ggcorrplot(cor_matrix, method = "circle", hc.order = TRUE, type = "upper",
           lab = TRUE, lab_size = 3)  # Enables labels and sets label size
```

## Hypothesis 1: Psychopathy and Exploitation in the PDG

To begin, I wanted to examine the association between psychopathy and exploitation (i.e., defection) in the PDG. I did this by using three different estimators of the relationship.

```{r psyc defect cor, echo=FALSE}

# Using the Percentage Bend Correlation Coefficient (see Wilcox, 2012) with 1000 Bootstrapped samples

pbcor(data$PDG_Avg, data$SRP_Total, ci = TRUE, nboot = 1000)

```

## Hypothesis 2: Psychopathy and Success in the PDG (Average Payoff)

Next, I wanted to examine the association between psychopathy and success (i.e., payoff) in the PDG. I did this by using three different estimators of the relationship between psychopathy and average payoff.

```{r psyc average payoff cor, echo=FALSE}

# Psyc-Avg Payoff using the Percentage Bend Correlation Coefficient (see Wilcox, 2012) with 1000 Bootstrapped samples

pbcor(data$PDGF_Avg, data$SRP_Total, ci = TRUE, nboot = 1000)

```

## Hypothesis 3: Psychopathy and Success in the PDG 

I also wanted to examine the association between psychopathy and success (i.e., payoff) in the PDG using relative payoffs.

```{r psyc rel payoff cor, echo=FALSE}

# Psyc-Rel Payoff using the Percentage Bend Correlation Coefficient (see Wilcox, 2012) with 1000 Bootstrapped samples

pbcor(data$Diff_Avg, data$SRP_Total, ci = TRUE, nboot = 1000)

```


## Hypothesis 4: Examining the Moderating Effect of Game Type on the Psychopathy-Exploitation Relationship

One of the central questions of this study was whether the game type (i.e., one-shot vs. iterated) moderated the association between psychopathy and exploitation. For this (and subsequent) analysis, the dataset was transformed to a long format for moderation analysis. A mixed effects model was used to account for the dependency between data points (i.e., via participants).

```{r psyc exploit mod, echo=FALSE}

# Select the specified columns

filtered_data <- data[, c("Participant_ID", "SRP_Total", "OS_Avg", "IT_Avg")]

# Transform data to long format

long_data <- pivot_longer(
  filtered_data,
  cols = c("OS_Avg", "IT_Avg"),
  names_to = "Game_Type",
  values_to = "Decision"
)

# Step 1: Convert 'Game_Type' to numeric where "OSF_Avg" is 0 and "ITF_Avg" is 1

long_data <- long_data %>%
  mutate(Game_Type = case_when(
    Game_Type == "OS_Avg" ~ 0,
    Game_Type == "IT_Avg" ~ 1,
    TRUE ~ as.integer(NA)))

# Step 2: Convert the numeric 'Game_Type' to a factor with labels

long_data <- long_data %>%
  mutate(Game_Type = factor(Game_Type, levels = c(0, 1), labels = c("One-shot", "Iterated")))

# Create linear mixed effects model

model <- lmer(Decision ~ SRP_Total + Game_Type + SRP_Total:Game_Type + (1 | Participant_ID), data = long_data, control=lmerControl(check.conv.singular = .makeCC(action = "ignore",  tol = 1e-4)))

# Create APA style table with additional parameters (see sjPlot documentation)

tab_model(
  model,
  show.std = "std"
)

# Examine marginal means 

test_predictions(model, c("SRP_Total", "Game_Type"), test = NULL)

plot_model(model, type = "int")

```


## Hypothesis 5: Examining the Moderating Effect of Game Type on the Psychopathy-Success Relationship (Average Payoff)

Another central question in this study was whether the association between psychopathy and success (as meaured via average payoff) varied as a function of game type (i.e., one-shot vs. iterated) produced significantly different associations between psychopathy and success (as measured via average payoff) Analysis was conducted in the same manner as Hypothesis 4.

```{r psyc average payoff mod, echo=FALSE}

# Select the specified columns

filtered_data <- data[, c("Participant_ID", "SRP_Total", "OSF_Avg", "ITF_Avg")]

# Transform data to long format

long_data <- pivot_longer(
  filtered_data,
  cols = c("OSF_Avg", "ITF_Avg"),
  names_to = "Game_Type",
  values_to = "Average_Payoff"
)

# Step 1: Convert 'Game_Type' to numeric where "OSF_Avg" is 0 and "ITF_Avg" is 1

long_data <- long_data %>%
  mutate(Game_Type = case_when(
    Game_Type == "OSF_Avg" ~ 0,
    Game_Type == "ITF_Avg" ~ 1,
    TRUE ~ as.integer(NA)))

# Step 2: Convert the numeric 'Game_Type' to a factor with labels

long_data <- long_data %>%
  mutate(Game_Type = factor(Game_Type, levels = c(0, 1), labels = c("One-shot", "Iterated")))

# Create linear mixed effects model

model <- lmer(Average_Payoff ~ SRP_Total + Game_Type + SRP_Total:Game_Type + (1 | Participant_ID), data = long_data, control=lmerControl(check.conv.singular = .makeCC(action = "ignore",  tol = 1e-4))
)

# Create APA style table with additional parameters (see sjPlot documentation)

tab_model(
  model,
  show.std = "std"
)

# Examine marginal means 

test_predictions(model, c("SRP_Total", "Game_Type"), test = NULL)

plot_model(model, type = "int")

```

## Hypothesis 6: Examining the Moderating Effect of Game Type on the Psychopathy-Success Relationship (Relative Payoff)

Lastly, the same moderation analysis from Hypothesis 5 was conducted with relative payoff as the outcome measure. 

```{r psyc relative payoff mod, echo=FALSE}

# Select the specified columns

filtered_data <- data[, c("Participant_ID", "SRP_Total", "Diff_OS_Avg", "Diff_IT_Avg")]

# Transform data to long format

long_data <- pivot_longer(
  filtered_data,
  cols = c("Diff_OS_Avg", "Diff_IT_Avg"),
  names_to = "Game_Type",
  values_to = "Relative_Payoff"
)

# Step 1: Convert 'Game_Type' to numeric where "OSF_Avg" is 0 and "ITF_Avg" is 1

long_data <- long_data %>%
  mutate(Game_Type = case_when(
    Game_Type == "Diff_OS_Avg" ~ 0,
    Game_Type == "Diff_IT_Avg" ~ 1,
    TRUE ~ as.integer(NA)))

# Step 2: Convert the numeric 'Game_Type' to a factor with labels

long_data <- long_data %>%
  mutate(Game_Type = factor(Game_Type, levels = c(0, 1), labels = c("One-shot", "Iterated")))

# Create linear mixed effects model

model <- lmer(Relative_Payoff ~ SRP_Total + Game_Type + SRP_Total:Game_Type + (1 | Participant_ID), data = long_data, control=lmerControl(check.conv.singular = .makeCC(action = "ignore",  tol = 1e-4))
)

# Create APA style table with additional parameters (see sjPlot documentation)

tab_model(
  model,
  show.std = "std"
)

# Examine marginal means 

test_predictions(model, c("SRP_Total", "Game_Type"), test = NULL)

plot_model(model, type = "int")

```